\documentclass[10pt,aspectratio=169]{beamer}
\usetheme{default}
\setbeamercovered{invisible}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{
    \flushright{\hfill \insertframenumber{}/\inserttotalframenumber}
}

\usepackage{listings}

% User-defined colors.
\definecolor{DarkGreen}{rgb}{0, .5, 0}
\definecolor{DarkBlue}{rgb}{0, 0, .5}
\definecolor{DarkRed}{rgb}{.5, 0, 0}
\definecolor{LightGray}{rgb}{.95, .95, .95}
\definecolor{White}{rgb}{1.0,1.0,1.0}
\definecolor{darkblue}{rgb}{0,0,0.9}
\definecolor{darkred}{rgb}{0.8,0,0}
\definecolor{darkgreen}{rgb}{0.0,0.85,0}

% Settings for listing class.
\lstset{
  language=C++,                        % The default language
  basicstyle=\small\ttfamily,          % The basic style
  backgroundcolor=\color{White},       % Set listing background
  keywordstyle=\color{DarkBlue}\bfseries, % Set keyword style
  commentstyle=\color{DarkGreen}\itshape, % Set comment style
  stringstyle=\color{DarkRed}, % Set string constant style
  extendedchars=true % Allow extended characters
  breaklines=true,
  basewidth={0.5em,0.4em},
  fontadjust=true,
  linewidth=\textwidth,
  breakatwhitespace=true,
  showstringspaces=false,
  lineskip=0ex, %  frame=single
}

% changing math font
\usefonttheme[onlymath]{serif}


\begin{document}
    \title{A Runge-Kutta-Fehlberg solver\protect\\using traits and concepts\protect\\(part I)}
    \author{Alberto Artoni}
    \date{19/04/2024}

\begin{frame}[plain, noframenumbering]
    \maketitle
\end{frame}

\begin{frame}{Runge Kutta methods}
Consider an ODE of the form
\begin{equation*}
	\frac{\mathrm{d} y}{\mathrm{d} t} = f\ (t, y).
\end{equation*}

Given the Butcher coefficients $b_i, a_{ij}, c_i$, a generic Runge Kutta method with $p$ steps has the following update rule:
\begin{equation*}
u_{n+1} = u_{n} + h_n \sum_{i=1}^p b_i k_i,
\end{equation*}
where $h_n = t_{n+1} - t_n$, and where
$	k_i = f\ (t_n + c_i h_n, \ y_n + h_n\sum_{j=1}^{i-1}a_{ij}k_j).
$
\end{frame}

\begin{frame}{Runge Kutta Fehlberg (RK45)}
	RKF methods try to adaptively choose the step size. The strategy computes a low order $y^*_{n+1}$ approximation and a high order approximation  $y_{n+1}$. \\
	If the difference between the two approximations is too large, we decrease the step size. \\
	If the difference between the two step size is too small, we increase the step size.
	\end{frame}
	
\begin{frame}{RKF - The Algorithm}
	
	Given the Butcher coefficients $b_i, b_i^*, c_i, a_{ij}$, and by denoting $h_n = t_{n+1} - t_n$, the RKF algorithm reads:
		
	\begin{enumerate}
		\item Compute $k_i = f\left(t_n + c_i h_n, y_n + h_n\sum_{j=1}^{i-1}a_{ij}k_j\right)$.
		\item Compute the high-order solution \(y_{n+1} = y_{n} + h_n \sum_{i=1}^p b_i k_i\).
		\item Compute the low-order solution \(y_{n+1}^* = y_{n} + h_n \sum_{i=1}^p b_i^* k_i\).
		\item Compute the error \(\varepsilon_{n+1} = y_{n+1} - y_{n+1}^* = h_n \sum_{i=1}^p(b_i - b_i^*)k_i\).
		\item Adapt the step size \(h_{n+1} = \tau_{n} h_{n}\), where \(\tau_{n}\) is a prescribed reduction (\(<1\))/expansion (\(>1\)) factor depending on whether \(\varepsilon_{n+1}\) is larger or smaller than a prescribed tolerance.
	\end{enumerate}
\end{frame}


\begin{frame}{RKF solver}
This exercise (a modified version of \texttt{Examples/src/RKFSolver}) is about a set of tools that implements embedded Runge-Kutta-Fehlberg (explicit) methods to solve non-linear scalar and vector Ordinary Differential Equations, based on the \texttt{Eigen} library.
\end{frame}

\begin{frame}{Exercise: RKF solver}
The code structure is the following:
\begin{itemize}
\item \texttt{ButcherRKF} contains the definition of the Butcher tableaux for some common RKF methods.
\item \texttt{RKFTraits} defines the basic structure that enables to bind the type of the equation(s) to be solved (\textit{i.e.} scalar or vector).
\item \texttt{RKFResult} is a data structure containing the output of the RKF solver.
\item \texttt{RKF} implements a generic RKF solver interface, filling a proper \texttt{RKFResult} object.
\end{itemize}
\end{frame}

\begin{frame}{Exercise: RKF solver}
Starting from the provided solution sketch:
\begin{enumerate}
\item Implement the \texttt{concept}(s) defining a scalar and a vector in \texttt{RKFTraits.hpp}.
\item Implement the \texttt{RKF::RKFstep} method, performing just one timestep of the RKF method.
\item Implement the \texttt{RKF::solve()} time-advancing method, without error correction.
\item Use the just implemented solver to solve the model problem and the Van der Pol oscillator problem defined in the \texttt{main.cpp} file.
\item Include error correction into the solver and compare the results.
\end{enumerate}
\end{frame}

\begin{frame}{Motivations: why do we 
	employ concepts?}
	Using \texttt{concepts} allows you to write more expressive and robust code by specifying requirements on template parameters.
	
	\begin{itemize}
		\item Improved Readability: Concepts help make your code more readable by explicitly stating the requirements of template parameters.
		\item Compile-Time Constraints: Concepts allow you to enforce constraints on template arguments at compile time. 
	\end{itemize}
	
\end{frame}

\begin{frame}{}
	Test the code to solve following problem:
	\begin{equation}
\frac{dy}{dt} = f,
	\end{equation}
	where
	\begin{itemize}
		\item Source term: $f = -10y$.
		\item Analytical solution: $y =
	 \exp(-10t)$
		\item $t_0 = 0$.
		\item $t_f = 10$.
		\item $y_0 = 1$.
	\end{itemize}
			
	For the numerical scheme, use $h_0 = 0.2$ and $\epsilon = 10^{-4}$.
\end{frame}


\begin{frame}{Additional work}
  Maybe you want to extend the class to include Diagonally Implicit Runge Kutta Methods
  (DIRK), where the $A$ matrix of the Butcher array satisfies $A_{ij}=0$ if $j<i$, but
  $A_{ii}$ may be different from $0$.
  \smallskip
  
  In this case, at each stage we need to solve a non-linear system. Try to investigate the changes needed to implement a DIRK starting from the already implemented RKF solution.
  \smallskip
  
  A possible implementation is in \texttt{Examples/src/RKFSolver/}.
\end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
